---
html_document:
  df_print: kable
author: "Wesley Barnes"
date: "Spring Semester 2023"
output:
  word_document:
  pdf_document: default
  html_document:
    df_print: paged
title: "Associations between parenting styles and depression"
pdf_document:
  df_print: kable
editor_options: 
  markdown: 
    wrap: 72
zotero: "HGAPS"
---

```{r setup}
library(tidyverse) #wrangling
library(haven) #importing SPSS and other data formats
library(ggplot2) #deluxe graphics
library(ggExtra) #extra fancy graphics
library(psych) #one package that covers ~80% of what we use in psychology 
library(rstatix) #a package that offers useful options for correlation tables, ANOVA
library(vcd) #visual displays of categorical data
library(mosaic) #more tools for categorical data
library(rockchalk) #useful for intermediate/advanced regression analyses
library(beeswarm)
library(decoder)
library(devtools)
library(ggcorrplot)
library(dplyr) # load dplyr
library(knitr)
options(digits = 3, scipen = NULL)
knitr::opts_chunk$set(echo = TRUE)
```

Note: v3.2 of the data had an error in item 10 (it was accidentally
over-written by the 10 item scale score)

```{r Load data}
# [] come back to this and load the datasets for class
ku2019<-read_spss("Korea_Spring_Fall_2017 v5 (safe)+2018+2019+Spr2020 (1).sav")
```


#Stuff for 525 ##Checking assumptions

```{r Checking assumptions; add gender/sex to subset}
modelvars<-ku2019 %>% select('Gender','Race', 'Age','gbi7up', 'apqMomInvolveP',	
                               'apqDadInvolveP', 'apqPosParP', 'apqPoorMonP', 'apqInconDiscP', 'apqCorpPunP', 'IAPfacAuthoritative', 'IAPfacAuthoritarian') #make data subset
describe(modelvars) #Get N, min & max (compare to source docs), skew & kurtosis
pairs.panels(modelvars) #Look at histograms, scatterplots, LOESS smoother
```

##Basic missing data handling -- listwise deletion Contrast this with
"pairwise," earlier.

```{r Listwise delete}
modelvars<-na.omit(modelvars)
describe(modelvars) #Get N, min & max (compare to source docs), skew & kurtosis
pairs.panels(modelvars) #Look at histograms, scatterplots, LOESS smoother
```

```{r  Descriptives}
get_summary_stats(modelvars) #gives IQR, but no skew or kurtosis

freq_table(modelvars, gbi7up) # Gives counts and bin percentages, not cumulative

mahalanobis_distance(modelvars) %>% arrange(desc(mahal.dist)) #very helpful for outliers

```
# Correlations

```{r Correlations}
cor.mat <- modelvars %>% cor_mat() #table of coefficients (Pearson as default)
cor.mat

cor.mat %>% cor_get_pval() #table of p values

cor.mat %>%
  cor_as_symbols() %>%
  pull_lower_triangle() #triangle of asterisks
```


# R and R squared value

```{r and R^2 value}

#r
cor(modelvars$apqDadInvolveP*modelvars$Race, modelvars$gbi7up)

# r squared
summary(setCor(gbi7up~apqDadInvolveP:Race, data=modelvars, std=TRUE))

```


# Total Correlation Plot

```{Total Correlation Plot}

ggcorrplot(cor(modelvars),
           hc.order = TRUE,
           type = "lower",
           outline.color = "white")
```


# Inconsistent Discipline and Depression

```{Inconsistent Discipline and Depression}
beeswarm(gbi7up ~ apqInconDiscP, data = modelvars,
         method ='swarm',
         pch =16, 
         col =c("#4B9CD3"),
         xlab ='Inconsistent Discipline', ylab ='Depression',
         main ="Inconsistent Discipline and Depression")
boxplot(gbi7up ~ apqInconDiscP, data = modelvars,add = T, col =c("#3256a822","#a8325822"), border ="black")
```


```{Poor Montioring and Depression}
beeswarm(gbi7up ~ apqPoorMonP, data = modelvars,
         method ='swarm',
         pch =16, 
         col =c("gold"),
         xlab ='Poor Monitoring', ylab ='Depression',
         main ="Poor Monitoring and Depression")
boxplot(gbi7up ~ apqPoorMonP, data = modelvars,add = T, col =c("#3256a822","#a8325822"), border ="black")

```



# Multiple Regression Including Race


```{Multiple Regression Including Race}


DepressionGraph = ggplot(modelvars, aes(x=gbi7up, y=apqDadInvolveP,colour=as.factor(Race)))+geom_point() + 
  scale_color_manual(values=c("#7BAFD4", "black",'#8D9092','gold')) + geom_smooth(method="lm", se=FALSE) +
  xlab("Paternal Involvement")+ ylab("Depression Score") +ggtitle("Paternal Involvement and Depression Scores, with Race")

ggMarginal(DepressionGraph, groupColour=TRUE, groupFill = TRUE)
DepressionGraph

```


# Paternal Involvement Correlations

```{r}

# normal paternal involvement correlative score
cor(modelvars$apqDadInvolveP, modelvars$gbi7up)

# isolated correlative score (African Americans)
AA = modelvars[modelvars$Race==3,]
cor(AA$apqDadInvolveP, AA$gbi7up)


```

```{r}
#write.csv(modelvars, "WesleyBarnesSufficientData.csv")
```




